Secure Composition of Robust and Optimising Compilers

### 2023-07-05 popl/main.tex
transition sentence from 2.1 to 2.2
Beginning of spec model

This section introduces a trace model (\Cref{subsec:propdefs}) that is used to define properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct}).
The properties are of practical importance and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents.
The study uses monitors, which are able to check wether a trace admits a property or not.
The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.
So, this section also presents monitors for each of the defined properties (\Cref{subsec:monitors}) and introduces monitor-based satisfaction.
Finally, for each property, it is shown that monitor-based satisfaction implies property satisfaction.

@@

Security tags describe wether an event contains sensitive information ($\lock$) or not ($\unlock$).
Control tags state wether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
Pre-events describe the actual kind of event that happened.
$\ev{Alloc\ \loc\ n}$ states that an allocation of size $n$ happened and is stored at $\loc$.
Locations $\loc$ are elements of $L$ and left abstract.
However, it is expected that $L$ is closed over addition with natural numbers.
$\ev{Dealloc\ \loc}$ announces that location $\loc$ is freed.
$\ev{Use\ \loc\ n}$ abstracts over reads from $\loc$ and writes to $\loc$ in memory, where $n$ is read/written, respectively.
$\ev{Branch\ n}$ and $\ev{Binop\ n}$ leak values involved in a branch or in a binary operation with data-dependent execution time.
Finally, events are either the emptyevent ($\emptyevent$), a program crash ($\lightning$), or a tuple consisting of a pre-event, a control tag, and a security tag.

This trace model is used to define any property of interest.
In a sense, the trace model is most general, because it abstracts all events done in the languages as defined in \Cref{sec:casestud}.
This papers uses $\event[_{1}]\le_{\trace}\event[_{2}]$ if $\event[_{1}]$ occurs at an earlier position in $\trace$ than $\event[_{2}]$.

>>>>

This section introduces a trace model (\Cref{subsec:propdefs}) that is used to define some example properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct}).
These properties are of practical importance and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents.
The study uses monitors, which are able to check wether a trace admits a property or not.
The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.
So, this section also presents monitors for each of the defined properties (\Cref{subsec:monitors}) and introduces monitor-based satisfaction.
Finally, for each property, it is shown that monitor-based satisfaction implies property satisfaction.

@@

The specification trace model has events that carry different meta-information, such as security- and control--tags.
Security tags describe wether an event contains sensitive information ($\lock$) or not ($\unlock$).
Control tags state wether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
%Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
Now, pre--events describe the actual kind of event that happened.
One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that happens whenever a program claims $n$ cells of memory and stores them at address $\loc$.
Likewise, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
These two events alone allow already a partial description of \gls{tms} by requiring that e.g. there is only one deallocation event that carries a location $\loc$.
To be able to express \gls{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
For cryptographic code, there is a general guideline that secrets must not be visible on a trace.
Moreover, an instruction whose timing is data--dependent must not have a secret as an operand.
Typical operations with data--dependent timing are branches and certain binary operations, such as division.
Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).
Lastly, events are either the empty event ($\emptyevent$), a program crash ($\lightning$), or a tuple consisting of a pre-event, a control--tag, and a security--tag.
